Physcraper: A Python package for continually updated phylogenetic trees using the Open Tree of Life

### 2021-04-20 sources/bmc_template_submission/bmc_article.tex
page 2 MK edits

Decades of single locus sequencing have generated
massive amounts of homologous DNA datasets that have the potential to be used for phylogenetic
reconstruction at many scales \cite{sanderson2008phylota}.
More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
million DNA sequences that were gauged to have the potential to resolve phylogenetic
relationships of 98.05\% of the almost 241,000
distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
However, even thirteen years later, phylogenetic estimates for most of these taxa
are still not available \cite{mctavish_how_2017}.
The Open Tree of Life project assembles a comprehensive synthetic tree of life comprising
2.3 million tips, of which around only
90,000 are supported by phylogenetic data - the remaining 1.4 million taxa are placed
in the tree based on taxonomy.
There is a considerable amount of phylogenetically
informative data in GenBank with the potential to fill these phylogenetic gaps in
the tree of life, but this data either has not been analysed or the analyses
have not been made publicly available \cite{mctavish_how_2017}.

@@

Taxon names in the input tree must be standardized to OpenTree taxonomy
\cite{rees2017automated} using OpenTree's bulk Taxonomic Name Resolution Service
(TNRS) tool \cite{TNRStool}. Users can upload their
own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's
Phylesystem \cite{phylesystemGithub, mctavish2015phylesystem} that also have
alignments available on TreeBASE \cite{piel2009treebase}.

>>>>

Decades of single locus sequencing have generated
massive amounts of homologous DNA datasets that have the potential to be used for phylogenetic
reconstruction at many scales \cite{sanderson2008phylota}.
More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
million DNA sequences that were gauged to have the potential to resolve phylogenetic
relationships of 98.05\% of the almost 241,000
distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
However, even thirteen years later, phylogenetic estimates for most of these taxa
are still not available \cite{mctavish_how_2017}.
The Open Tree of Life project assembles a comprehensive synthetic tree of life comprising
2.3 million tips, of which around only
90,000 are supported by phylogenetic data uploaded to OpenTrees' database (the Phylesystem) by curators - the remaining 1.4 million taxa are placed
in the tree based on taxonomy.
There is a considerable amount of phylogenetically
informative data in GenBank with the potential to fill these phylogenetic gaps in
the tree of life, but this data either has not been analysed or the analyses
have not been made publicly available \cite{mctavish_how_2017}.

@@

Taxon names in the input tree must be standardized to OpenTree taxonomy
\cite{rees2017automated} using OpenTree's bulk Taxonomic Name Resolution Service
(TNRS) tool \cite{TNRStool}. Users can upload their
own tree, or choose from among the 2, 950 curated trees stored in OpenTree's
Phylesystem database \cite{phylesystemGithub, mctavish2015phylesystem} that also have
alignments available on TreeBASE \cite{piel2009treebase}.
